---
title: "Assignments"
description: |
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

As you go, you may develop your own R Markdown style you like or you might use my file layout to start. If you'd like a different starting point, you can use the template below.

```{r}
library(downloadthis)

download_file(
  path = "assignments/template.Rmd",
  button_label = "Download template.Rmd file",
  button_type = "warning",
  has_icon = TRUE,
  icon = "fa fa-save",
  self_contained = FALSE
)
```

### Fall 2021 Assignments

#### Assignment #1

Goals:

* Set up and use git and GitHub with RStudio.  
* Create a personal website and hear about why this is a useful thing to have.  
* Start generating ideas about your project for the course.
* Read about and reflect on bias and fairness in data science.

```{r}
download_file(
  path = "assignments/2021_fall/01_assignment.html",
  button_label = "Download Assignment #1!",
  button_type = "warning",
  has_icon = TRUE,
  icon = "fa fa-save",
  self_contained = FALSE
)
```

#### Assignment #2

Goals:

* Continue using and become more comfortable with Git and GitHub.  
* Refresh your machine learning skills.  
* Use the `tidymodels` packages to build models.  
* Read about and reflect on bias and fairness in data science.


```{r}
download_file(
  path = "assignments/2021_fall/02_assignment.html",
  button_label = "Download Assignment #2!",
  button_type = "warning",
  has_icon = TRUE,
  icon = "fa fa-save",
  self_contained = FALSE
)
```

#### Assignment #3

Goals:

* Continue using and become more comfortable with Git and GitHub.  
* Build more machine learning models using `tidymodels`, including `xgboost`.    
* Build and use a stacked model.  
* Start thinking about a shiny app.  
* Explore new packages from "Function Friday".
* Continue learning about bias and fairness in data science.  

```{r}
download_file(
  path = "assignments/2021_fall/03_assignment.html",
  button_label = "Download Assignment #3!",
  button_type = "warning",
  has_icon = TRUE,
  icon = "fa fa-save",
  self_contained = FALSE
)
```

#### Assignment #4

Goals:

* Continue using and become more comfortable with Git and GitHub.  
* Use SQL inside of R.  
* Make some amazing graphs!  
* Create a shiny app using the slightly more advanced Shiny tips we covered in class.    
* Explore new packages from "Function Friday".

```{r}
download_file(
  path = "assignments/2021_fall/04_assignment.html",
  button_label = "Download Assignment #4!",
  button_type = "warning",
  has_icon = TRUE,
  icon = "fa fa-save",
  self_contained = FALSE
)
```

### Spring 2021 Assignments


#### Assignment #1

Goals:

* Set up and use git and GitHub with RStudio.  
* Create a personal website and be motivated to do it.  
* Create a small package (optional).  
* Refresh your machine learning skills.  
* Use the `tidymodels` packages to build models.  
* Read about and reflect on bias and fairness in data science.

```{r}
download_file(
  path = "assignments/01_assignment.html",
  button_label = "Download Assignment #1!",
  button_type = "warning",
  has_icon = TRUE,
  icon = "fa fa-save",
  self_contained = FALSE
)
```

#### Assignment #2

Goals:

* Continue using and become more comfortable with Git and GitHub.  
* Build more machine learning models using `tidymodels`.  
* Build and use a stacked model.  
* Use some global model interpretation tools.  
* Create a shiny app that allows you to interact with a model.  
* Add some customized theming to a shiny app.  
* Continue learning about bias and fairness in data science.  


```{r}
download_file(
  path = "assignments/02_assignment.html",
  button_label = "Download Assignment #2!",
  button_type = "warning",
  has_icon = TRUE,
  icon = "fa fa-save",
  self_contained = FALSE
)
```


#### Assignment #3

Goals:

* Continue using and become more comfortable with Git and GitHub.  
* Use some local model interpretation tools.  
* Use SQL inside of R.  
* Make some amazing graphs!  
* Practice using `geom_sf()`  
* Practice using `tidytext`  
* Continue learning about bias and fairness in data science.  


```{r}
download_file(
  path = "assignments/03_assignment.html",
  button_label = "Download Assignment #3!",
  button_type = "warning",
  has_icon = TRUE,
  icon = "fa fa-save",
  self_contained = FALSE
)
```

#### Assignment #4

Goals:

* Practice using `tmap`  
* Practice using regular expressions via `stringr`  
* Continue learning about bias and fairness in data science.  


```{r}
download_file(
  path = "assignments/04_assignment.html",
  button_label = "Download Assignment #4!",
  button_type = "warning",
  has_icon = TRUE,
  icon = "fa fa-save",
  self_contained = FALSE
)
```


#### Assignment #5

* Practice using `countrycode` R package.

```{r}
download_file(
  path = "assignments/05_assignment.html",
  button_label = "Download Assignment #5!",
  button_type = "warning",
  has_icon = TRUE,
  icon = "fa fa-save",
  self_contained = FALSE
)
```


