---
title: "Assignments"
description: |
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

As you go, you may develop your own R Markdown style you like. I encourage you to download and use the template below at first.

```{r}
library(downloadthis)

download_file(
  path = "assignments/template.Rmd",
  button_label = "Download template.Rmd file",
  button_type = "warning",
  has_icon = TRUE,
  icon = "fa fa-save",
  self_contained = FALSE
)
```


## Assignment #1

Goals:

* Set up and use git and GitHub with RStudio.  
* Create a personal website and be motivated to do it.  
* Create a small package (optional).  
* Refresh your machine learning skills.  
* Use the `tidymodels` packages to build models.  
* Read about and reflect on bias and fairness in data science.

```{r}
download_file(
  path = "assignments/01_assignment.html",
  button_label = "Download Assignment #1!",
  button_type = "warning",
  has_icon = TRUE,
  icon = "fa fa-save",
  self_contained = FALSE
)
```

## Assignment #2

Goals:

* Continue using and become more comfortable with Git and GitHub.  
* Build more machine learning models using `tidymodels`.  
* Build and use a stacked model.  
* Use some global model interpretation tools.  
* Create a shiny app that allows you to interact with a model.  
* Add some customized theming to a shiny app.  
* Continue learning about bias and fairness in data science.  


```{r}
download_file(
  path = "assignments/02_assignment.html",
  button_label = "Download Assignment #2!",
  button_type = "warning",
  has_icon = TRUE,
  icon = "fa fa-save",
  self_contained = FALSE
)
```
