---
title: 'Assignment #4'
output: 
  html_document:
    toc: true
    toc_float: true
    df_print: paged
    code_download: true
---

```{r setup, include=FALSE}
#knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)
```

```{r libraries, message=FALSE}
#Regular expressions
library(tidyverse)        # contains stringr for regex
library(googlesheets4)    # for reading in data
gs4_deauth()              # to read in google sheet (or download)

#tmap

theme_set(theme_minimal()) # Lisa's favorite theme
```

When you finish the assignment, remove the `#` from the options chunk at the top, so that messages and warnings aren't printed. If you are getting errors in your code, add `error = TRUE` so that the file knits. I would recommend not removing the `#` until you are completely finished.

## Put it on GitHub!        

From now on, GitHub should be part of your routine when doing assignments. I recommend making it part of your process anytime you are working in R, but I'll make you show it's part of your process for assignments.

**Task**: When you are finished with the assignment, post a link below to the GitHub repo for the assignment. If you want to post it to your personal website, that's ok (not required). Make sure the link goes to a spot in the repo where I can easily find this assignment. For example, if you have a website with a blog and post the assignment as a blog post, link to the post's folder in the repo. As an example, I've linked to my GitHub stacking material [here](https://github.com/llendway/ads_website/tree/master/_posts/2021-03-22-stacking).

## Regular Expressions

Either read in the data using the code below (need to install `googlesheets4` library if you don't have it) or download from [this](
https://drive.google.com/file/d/12oqEt11miNGL_MtIcU75jx84pz6VEq_w/view?usp=sharing) URL and put it in your project folder.

```{r}
bestsellers <- read_sheet("https://docs.google.com/spreadsheets/d/1n3xKHK4-t5S73LgxOJVJWT5fMYjLj7kqmYl1LHkpk80/edit?usp=sharing")
```

**BE SURE TO REMOVE eval=FALSE** from all code sections.

1. Find books with multiple authors (HINT: Consider the possibility of an author having “and” in their name)

```{r, eval=FALSE}
bestsellers %>% 
  mutate(mult_auth =str_???(???, "???")) %>% 
  filter(mult_auth)
```

2. Detect if the author’s first name starts with a vowel

```{r, eval=FALSE}
bestsellers %>% 
  mutate(vowel =str_???(author, "???")) %>% 
  filter(vowel)
```

3. Change all authors with the name Alexander to Alex

```{r, eval=FALSE}
bestsellers %>% 
  mutate(names = str_???_???(???, "Alexander", "Alex"))
```

4. Find books that are the second book in a series

```{r, eval=FALSE}
bestsellers %>%
  mutate(series = str_???(???, 
                          pattern = "???")) %>%
  filter(series == TRUE)
```

5. Find books that are the third or fourth one in a series

```{r, eval=FALSE}
bestsellers %>%
  mutate(series = str_???(???, 
                          pattern = "???")) %>%
  filter(series == TRUE)
```

6. Find books that are the 10th, 11th, ..., or 19th book in a series

```{r, eval=FALSE}
bestsellers %>%
  mutate(series = str_???(???, 
                          pattern = "???")) %>%
  filter(series == TRUE)
```

7. Describe in your own words how you would go about writing a regular expression for password pattern matching (ie 8 character minimum, one capital letter minimum, one lowercase letter minimum, one digit minimum, one special character minimum)

## Data Ethics

**Task:**


